Mixed Non-Bayesian and DeGroot Learning

### 2012-12-11 docs/paper.tex
Added figure for trapezoidal distribution
Fixed image to work for both latex and pdflatex
Added some discussion of simulation and conclusion to paper.

\documentclass[a4paper,12pt]{article}
\usepackage{amsmath} % For 'cases' curly brace
\usepackage{amsfonts} % For bold number set font
\begin{document}
\title{Mixed Non-Bayesian and DeGroot Learning}
\author{Pedro d'Aquino\\pdaquino@umich.edu \and Dan Clark\\ddclark@umich.edu}
\date{December 11, 2012}

@@

The model allows for a variety of generative methods to be used for creation of the network structure such as Erdos-Renyi, Watts-Strogatz, Kleinberg's Small-World Model, or a power law distribution.  We investigate each of these network types in section \{ \}.% @todo[DDC] Refer to simulation results

@@

% @todo[DDC] Graph/figure of this would probably be good to have.

@@

@@

The second method we consider for creating a probability distribution over the set of belief states $\Theta$ is to define an exponential distance metric such that

\begin{equation}
\mu_{j,t}^\prime(\theta_k) \propto e^{-(m_{j,t}-\hat{m_k})^2}
\end{equation}
where $\mu_{j,t}^\prime$ is the belief distribution that is used for the DeGroot node $j$'s values of $\mu_{j,t}(\theta_k)$ in the Non-Bayesian update of equation~\ref{eq:non_bayesian_update}.

@@

\begin{equation}
q_{j,t} = \sum_{k=1}^n e^{-(m_{j,t}-\hat{m_k})^2}
\end{equation}

@@

\begin{equation}
\mu_{j,t}^\prime(\theta_k) = e^{-(m_{j,t}-\hat{m_i})^2} / q_{j,t}
\end{equation}

@@

@@

% Can discuss in conclusion that future work could either focus on figuring out how to do this or on developing other ways of doing the DeGroot-to-Non-Bayesian conversion that make such proofs easier.

>>>>

\documentclass[a4paper,12pt]{article}
\usepackage{amsmath} % For 'cases' curly brace
\usepackage{amsfonts} % For bold number set font
\usepackage{graphicx}
\usepackage[font=bf]{caption}
\begin{document}
\title{Mixed Non-Bayesian and DeGroot Learning}
\author{Pedro d'Aquino\\pdaquino@umich.edu \and Dan Clark\\ddclark@umich.edu}
\date{December 11, 2012}

@@

The model allows for a variety of generative methods to be used for creation of the network structure such as Erdos-Renyi, Watts-Strogatz, Kleinberg's Small-World Model, or a power law distribution.  We investigate several of these network types in simulation, described in section~\ref{sec:simulation}.

@@

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{trapezoidalDistribution}
\caption{Trapezoidal Probability Distribution}
\end{figure}

@@

The second method we consider for creating a probability distribution over the set of belief states $\Theta$ is to define a distance metric such that

\begin{equation}
\mu_{j,t}^\prime(\theta_k) \propto \frac{1}{(m_{j,t}-\hat{m_k})^2}
\end{equation}
where $\mu_{j,t}^\prime$ is the belief distribution that is used for the DeGroot node $j$'s values of $\mu_{j,t}(\theta_k)$ in the Non-Bayesian update of equation~\ref{eq:non_bayesian_update}.

@@

\begin{equation}
q_{j,t} = \sum_{k=1}^n \frac{1}{(m_{j,t}-\hat{m_k})^2}
\end{equation}

@@

\begin{equation}
\mu_{j,t}^\prime(\theta_k) = \frac{(m_{j,t}-\hat{m_i})^{-2}}{q_{j,t}}
\end{equation}

@@

\section{Simulation}

In order to seek insight about the behavior of our model we performed a series of simulations using the JUNG network/graph software library \cite{website:JUNGGraphLibrary}.  For each simulation, $\Theta$ consisted of two Gaussian distributions with means of 100 and 105 and standard deviation of 1.  Each trial was run until all nodes have converged upon the true state or some maximum number of timesteps had elapsed.

\subsection{Node Ratio vs. Convergence Time}
\label{sec:simulation}

In order to characterize the effect of naive DeGroot nodes on Non-Bayesian social learning we simulated time taken for convergence versus the ratio of Non-Bayesian to DeGroot nodes in a network.  To simulate the spread of information throughout the network, only 50\% of the Non-Bayesian received any signals; the other 50\% needed to rely completely on learning from their neighbors.  Each trial was run 100 times for each ratio percentage between from 1-100\%, for a total of 10,000 trials for each graph type.

% Graph here

The data show a somewhat counterintuitive result: the presence of naive DeGroot nodes actually tends to accelerate the speed of convergence.  Further investigation showed this to be a result of the process used to convert DeGroot belief to Non-Bayesian belief.  The conversion method of section~\ref{sec:likelihood_metric} tends to result in a belief distribution that is heavily skewed to a single $\theta_i$.  This relative ``certainty'' of the DeGroot nodes is picked up by their Non-Bayesian neighbors, bringing them more quickly to convergence.

% Graph here

Thus we see that the presence of naive, non-observant agents in a graph can actually accelerate the rate of learning due to eagerness of these nodes to settle on a single set of beliefs.

\subsection{Graph Size vs. Convergence Time}

We also evaluated time to convergence versus graph size for several network types.  These simulations were performed with a ratio of 0.8 Non-Bayesian to DeGroot nodes, where 60\% of Non-Bayesian nodes received signals from the true probability distribution and 40\% had to rely completely on information from neighbors.

% Graph here

Time required for convergence appears to be linear with respect to the number of nodes in the graph.

@@

We have developed a model that allows the Non-Bayesian social learning method of Jadbabaie et. al. to take place in a graph that also contains nodes performing simple DeGroot-style belief updates.  This model can be used as a theoretical framework for evaluating the effect that naive, non-observant agents can have on a network of more sophisticated learners.

We performed a series of simulations to obtain greater intuition about our model.  Some results were counterintuitive, such as the fact that DeGroot nodes can actually accelerate convergence due to a tendency towards certainty in their beliefs.

Future work could focus on proving further theoretical results for our model, such as specific conditions necessary to guarantee convergence.  This may involve development of different methods for converting DeGroot to Non-Bayesian belief that allow for easier theoretical analysis.  Future work could also investigate the effects of different choices for this method on convergence time.
